---
title: "04-home-field-advantage"
output:
  html_document:
    theme: cosmo
    toc: yes
    toc_depth: 5
    toc_float: true
editor_options:
  chunk_output_type: inline
author: Shaswat Rajput
---

```{r ,echo=TRUE, include=FALSE}
library(dplyr)
library(ggplot2)
library(magrittr)
```
### World Series : 

The World Series is the annual championship series of Major League Baseball (MLB) in North America, contested since 1903 between the American League (AL) champion team and the National League (NL) champion team. The winner of the World Series championship is determined through a best-of-seven playoff, and the winning team is awarded the Commissioner's Trophy. As the series is played during the fall season in North America, it is sometimes referred to as the Fall Classic. Since 2017, it has been officially known as the World Series <br>

#### What is Home field Advantage ?
In team sports, the term home advantage – also called home ground, home field, home-field advantage, home court, home-court advantage, defender's advantage or home-ice advantage – describes the benefit that the home team is said to gain over the visiting team. This benefit has been attributed to psychological effects supporting fans have on the competitors or referees; to psychological or physiological advantages of playing near home in familiar situations; to the disadvantages away teams suffer from changing time zones or climates, or from the rigors of travel; and in some sports, to specific rules that favor the home team directly or indirectly. In baseball, in particular, the difference may also be the result of the home team having been assembled to take advantage of the idiosyncrasies of the home ballpark, such as the distances to the outfield walls; most other sports are played in standardized venues.

## If home field advantage exists, how much of an impact does it have on winning the world series? <br>

### SETUP :
<li> Suppose that the Braves and the Yankees are teams competing in the World Series.</li>
<li> The table below has the two possible schedules for each game of the series. (NYC = New York City, ATL = Atlanta) </li><br>

| Overall advantage | Game 1 | Game 2 | Game 3 | Game 4 | Game 5 | Game 6 | Game 7 |
|:-----------------:|:------:|:------:|:------:|:------:|:------:|:------:|:------:|
|       Braves      |   ATL  |   ATL  |   NYC  |   NYC  |   NYC  |   ATL  |   ATL  |
|      Yankees      |   NYC  |   NYC  |   ATL  |   ATL  |   ATL  |   NYC  |   NYC  |

<li> Let *P*<sub>*B*</sub> be the probability that the Braves win a
    single head-to-head match-up with the Yankees, under the assumption
    that home field advantage doesn’t exist. Let
    *P*<sub>*B*</sub><sup>*H*</sup> denote the probability that the
    Braves win a single head-to-head match-up with the Yankees as the
    home team (H for home). Let *P*<sub>*B*</sub><sup>*A*</sup> denote
    the probability that the Braves win a single head-to-head match-up
    with the away team (A for away).</li>
| Game location |    No advantage   | Advantage                                                            |
|:-------------:|:-----------------:|:---------------------------------------------------------------------|
|      ATL      | *P*<sub>*B*</sub> | *P*<sub>*B*</sub><sup>*H*</sup> = *P*<sub>*B*</sub> \* 1.1           |
|      NYC      | *P*<sub>*B*</sub> | *P*<sub>*B*</sub><sup>*A*</sup> = 1 − (1 − *P*<sub>*B*</sub>) \* 1.1 |
<br>

We are calculating the the probability that Braves win the World Series when the sequence of games location is {NYC, NYC, ATL, ATL, ATL, NYC, NYC}. <br>

### Setting the parameters for simulation and analytics answer. W
```{r}
require(dplyr)
require(data.table)
# Get all possible outcomes
apo <- fread("all-possible-world-series-outcomes.csv")
set.seed(44)
hfi <- c(0,0,1,1,1,0,0) # {NYC, NYC, ATL, ATL, ATL, NYC, NYC}
pb <- 0.55
advantage_multiplier <- 1.1
pbh <- 0.55*advantage_multiplier
pba <- 1 - (1 - 0.55)*advantage_multiplier

```

## ANALYTICAL:
### Probability that Braves win world Series without Home advantage. For this part we just simply used the pbinom functiom and set lower.tail to be False as we want the right side of the curve.
```{r}
# Without Home Advantage
analy_without<- pbinom(3,7,pb,lower.tail = F)
analy_without
```
So, we found that probability that Braves win world series without Home Field advantage is equal to 0.6082. This method was analytical method.<br>

Let's find out the probability that Braves win the world Series with Home Field Advantage using analytical method.<br>

```{r}
# With home Aadvantage 
apo[, p := NA_real_] # Initialize new column in apo to store prob
for(i in 1:nrow(apo)){
  prob_game <- rep(1, 7)
  for(j in 1:7){
    p_win <- ifelse(hfi[j], pbh, pba)
    prob_game[j] <- case_when(
        apo[i,j,with=FALSE] == "W" ~ p_win
      , apo[i,j,with=FALSE] == "L" ~ 1 - p_win
      , TRUE ~ 1
    )
  }
  apo[i, p := prod(prob_game)] # Data.table syntax
}

# Sanity check: does sum(p) == 1?
apo[, sum(p)] # This is data.table notation

# Probability of overall World Series outcomes
apo[, sum(p), overall_outcome]
```

So, we found out that the probability of Braves winning the World Series is 0.604221. We get this answer using the analytical method.

Now, let's compare our results of without and with home advantage. We can do this by relative and absolute error.<br>

There is difference in the analytical method if the team has advantage or without it. <br>

Absolute Error = 0.6082878 - 0.604221	= 0.0040668<br>
Relative error = 0.0040668/0.6082878 = 0.00668565<br>

| Method | Home Advantage | Probability | 
|:-----------------:|:------:|:------:|
|       Analytic      |  No  |   0.6082878  |
|      Analytic      |   Yes  | 0.604221 | 

<br>

| Error Type | Error |
|:-----------------:|:------:|
|       Absolute Error      |  0.004066  |  
|     Relative Error    |  0.00668565  | ]

So, we found out that with analytical solution there is not much difference with home or not home field advantage.<br>

Let's do the same thing as above using the simulation and comapre our results with analytical solution.<br>

## SIMULATION: 

Let's take out out size of 10000000 and take mean of when the output>3.

```{r}
# Without home advantage 
without_home_adv<-mean(rbinom(10000000,7,pb)>3)
without_home_adv
```
So, if we do simulation to find the probability of Braves winning without Home Field Advantage, we get probability of 0.608

Let's find probability of Braves winning world series with Home field Advantage. 

```{r}
#With Home Advantage
with_adv <- mean((rbinom(100000,4,pba)+ rbinom(100000,3,pbh))>3)
with_adv
```
So, from the simulation we found out that probability that Braves win World Series with home advantage is 0.6057.<br>

Let's explore the difference between analytical and simulation result.<br>

#### Difference Between Analytical and Simualtion Results.

Let's find out the Mean absolute error between analytical and simulated probability with Home field Advantage.<br>
```{r}
# Mean absolute Error (with home advantage)
(0.604221	 - with_adv)
```
We found out that the difference between simulated and analytical result is very small. The difference turned out to be 0.000581 which is nearly equal to 0. <br>

Let's find the relative error between analytical and simulated result:

```{r}
# Relative Error (with home advantage)
(0.604221	 - with_adv)/0.604221
```

The relative error between simulated and analytical solution turned out to be very neglible, nearly equal to 0. <br>

Let's explore the differene simulated and analytical solution of Braves winning the World Series without the home-field advantage : <br>
```{r}
#Mean absolute Error (without home advantage)
analy_without - without_home_adv
```
 Just like the result we got in with home-field advantage, we see there is not much difference between simulated and analytical solution for Braves winning world series without home-field advantage.<br>
We can expect the relative error to be neglible too.<br>

Let's find the relative error without home field advantage :
```{r}
## Relative Error (without home advantage)
(analy_without-without_home_adv)/analy_without
```
We found that relative error is very small, almost negligible. So, we can say that analyical and simulated results are almost same for Braves winning world Series without home-field advantage.


| Home Field Advantage | Relative Error | Absolute Error | 
|:-----------------:|:------:|:------:|:------:|
|       Yes      |       0.0009615687  |   0.000581  |  
|      No      |    0.0004484668  |   0.0002727969  |  

## Conclusion:
We found out that there is not much difference between Analytical and Simulated Result for Braves winning World Series. We also infer that having home-field advantage doesn't increase your chance of winning World Series according to the parameters we have set initially.<br>
Overall we can say that Braves have winning chance of ~60% of winning the world series, but this  is based on our initial parameters we have set. To get practical result we need past data of Braves winning matches against Yankees so that we can get actual winning probability and then we can do peoper analysis of Braves winning World Series.<br>



**References and Further Study:**
[Historical Trends in Home-Field Advantage by Cyril Morong](https://web.archive.org/web/20050330200457/http://www.geocities.com/cyrilmorong@sbcglobal.net/HomeRoad.htm)


























