---
title: "03-Discrete-Probability-Calculations"
author: "Shaswat Rajput"
output: html_notebook
---

### <b><u> What is Probability ? </u></b>
Probability is a branch of mathematics that deals with calculating the likelihood of a given event's occurrence, which is expressed as a number between 1 and 0. An event with a probability of 1 can be considered a certainty: for example, the probability of a coin toss resulting in either "heads" or "tails" is 1, because there are no other options, assuming the coin lands flat. An event with a probability of .5 can be considered to have equal odds of occurring or not occurring: for example, the probability of a coin toss resulting in "heads" is .5, because the toss is equally as likely to result in "tails." An event with a probability of 0 can be considered an impossibility: for example, the probability that the coin will land (flat) without either side facing up is 0, because either "heads" or "tails" must be facing up. A little paradoxical, probability theory applies precise calculations to quantify uncertain measures of random events.<br>

<b> What is a probabilistic model? </b><br>
A probabilistic model is a quantative description of a situation, a phenomenon or an experiment where outcome is uncertain.

#### <b><u> Axioms: </u></b><br>
The rules that any probabilistic model should satisfy are called axioms of probability.
<li> <b><u>Non-Negativity: P(A) > 0 </li>
<li> Normalisation: P(Ω) = 1</li>
<li> Additivity (Finite) : If two events are disjoint then, P(A ∪ B) = P(A) + P(B) </li></u></b>
These 3 axioms are the only requirement in order to have a legitimate probability model. <br>

#### <b><u>Role of Probability Theory: </u></b>
<li>A Framework for analysing phenomena with uncertain outcomes. </li>
<li>Rule for consistent reasoning. </li>
<li>Used for predictions and decisions. </li>

#### <b><u>Discrete Uniform Law: </u></b>
<li> Assume Ω consists of n equally likely elements. </li>
<li> Assume A consists of k elements. </li>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;So, P(A) = k/n

#### <b><u>Probability Calculation Steps: </u></b>
<li>Specify the sample space </li>
<li>Specify the probability law </li>
<li>Identify the event of interest </li>
<li>Calculate... </li>

### <b><u>Let's understand about Discrete Probability through following Problem: </u></b>
#### <b><u>How often does the better team win the World Series? </u></b><br>
The world series is a best-of-7 match-up between the champions of the American and National Leagues of Major League Baseball. In this assignment, you are going to explain probability calculations related to the world series. <br>

##### <b><u>Setup: </u></b><br>
<li>Suppose that the Braves and the Yankees are teams competing in the World Series. </li>
<li>Suppose that in any given game, the probability that the Braves win is P<sub>b</sub> and the probability that the Yankees win is P<sub>y</sub> = 1−P<sub>b</sub>. </li> <br>

<b><u>1. What is the probability that the Braves win the World Series given that P<sub>b</sub> = 0.55? </u></b><br>
```{r}
dbinom(4,7,0.55) + dbinom(5,7,0.55) +dbinom(6,7,0.55) +dbinom(7,7,0.55) 
pbinom(3,7,0.55,lower.tail = FALSE)
```
We can say that the probability of Braves winining the world series is ---->
P(WS) = P(WS in 4) + P(WS in 5) + P(WS in 6) + P(WS in 7) = 0.608<br>

<b><u>2. What is the probability that the Braves win the World Series given that P<sub>b</sub> = x? This will be a figure (see below) with PB on the x-axis and P(Braves win World Series) on the y-axis. </u></b><br>
```{r}
probab <- seq(0.5,1.0, length.out = 100)
win_probab <- rep(NA,100)
for (i in seq_along(probab)){
  win_probab[i] = pbinom(3,7,probab[i],lower.tail = FALSE)
}
plot(probab,win_probab,xlab = "Probability of Braves winning a head-to-head matchup",ylab = "P(Win World Series)")
```

<b><u>3. Suppose one could change the World Series to be best-of-9 or some other best-of-X series. What is the shortest series length so that P(Braves win World Series|Psub>b</sub> = 0.55) ≥ 0.8 ?</u></b>
Let's assume Braves are playing thre best of 11 series. In any given series, a minimum of more than half is required, so in this cade a minimum of 6 wins are required to win the World Series.
```{r}
ev_od <- function (x) {
  if (x %% 2 == 0) {
    return("even")
  } else {
    return ('odd')
  }
}

for (i in 3:100) {
  if (ev_od(i) == "even") {
    win_probab = pbinom((i / 2), i, 0.55, lower.tail = FALSE)
    if (win_probab >= 0.8) {
      print(i)
      break
    }
  } else if (ev_od(i) == "odd") {
    win_probab = pbinom(floor(i / 2), i, 0.55, lower.tail = FALSE)
    if (win_probab >= 0.8) {
      print(i)
      break
    }
  }
}

```

<b><u>4. What is the shortest series length so that P(Braves win World Series| P<sub>b</sub> = x) ≥ 0.8? This will be a figure (see below) with P<sub>b</sub> on the x-axis and series length is the y-axis. </u></b>

```{r}
# min_game <- vector()
# 
# for (i in 3:100){
#   for (j in seq_along(probab)){
#     if (pbinom(floor(i / 2), i, probab[j], lower.tail = FALSE) >= 0.8){
#       min_game <- c(min_game,i)
#     }
#   }
# }
# plot(x=probab,y=min_game, type = 'l')


p_winning <- function(req_wins,p){
  losses <- req_wins-1
  p_ws <-0
  
  while (losses >= 0 ){
    p_ws <- p_ws + dbinom(losses, req_wins,p)
    losses = losses-1
  }
  return(p_ws)
}

games <- seq(3,1500,2)
probab_b <- seq(0.51,1.0,0.0001)
min_win <- rep(NA,length(probab_b))

for(i in seq_along(probab_b)){
  for (j in seq_along(games)){
    req_wins <- floor(0.5*games[j]) +1
    probab_B_WS <- p_winning(req_wins = req_wins, probab_b[i])
    if(probab_B_WS >= 0.8){
      break
    }
  }
  min_win[i] <- games[j]
}

plot(y = min_win, x= rev(probab_b), type='l',lwd = 2.5,xlab = "Probability of Braves winning a head to head matchup", ylab = "Series Length")
```



<b><u>5. Calculate P(P<sub>b</sub> = 0.55|Braves win World Series in 7 games) under the assumption that either P<sub>b</sub> = 0.55 or P<sub>b</sub> = 0.45. Explain your solution.<br></u></b>
<li> Let's assume that the P(P<sub>b</sub> = 0.55) and P(P<sub>b</sub> = 0.45) = 0.5, meaning that the events are equally likely.</li>
```{r}
(0.5 * dnbinom(3,4,0.55)) / ((0.5 * dnbinom(3,4,0.55))+ (0.5 * dnbinom(3,4,0.45)) )
```
So, P(P<sub>b</sub> = 0.55 | Braves win World Series in 7 games) = P(P<sub>b</sub> = 0.55 ∩ Braves win World Series in 7 games) / P(B) <br>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;OR <br>
= P(B) * P(Braves win World Series in 7 games | P<sub>b</sub> = 0.55) / ((P(B) * P(Braves win World Series in 7 games | P<sub>b</sub> = 0.55) + (P(B) * P(Braves win World Series in 7 games | P<sub>b</sub> = 0.55)) = 0.55






