---
title: "Finale"
output: html_notebook
author: "Shaswat Rajput"
---


 Name: Shaswat Rajput
 
```{r}
library(stats)
library(dplyr)
library(statsr)
```


# 1. Simulation

The Monte Hall problem is a classic game show.  Contestants on the show where shown three doors.  Behind one randomly selected door was a sportscar; behind the other doors were goats.

At the start of the game, contestants would select a door, say door A.  Then, the host would open either door B or C to reveal a goat.  At that point in the game, the host would ask the contestant if she would like to change her door selection.  Once a contestant decided to stay or change, the host would open the choosen door to reveal the game prize, either a goat or a car.

In this problem, consider a **modified** version of the Monte Hall problem in which the number of doors is **variable**.  Rather than 3 doors, consider a game with 4 or 5 or 50 doors.  In the modified version of the game, a contestant would select an initial door, say door A.  Then, the host would open **one** of the remaining doors to reveal a goat.  At that point in the game, the host would ask the contestant if she would like to change her door selection.  Once a contestant decided to stay or change, the host would open the choosen door to reveal the game prize, either a goat or a car.

Consider two strategies:
  
  1. Always stay with the first door selected.
  2. Always switch to the unopened door.

**C.** The function `game` below plays a single game of Monte Hall.  The function returns a vector of length two, the first element is the prize under strategy 1 and the second element is the prize under strategy 2.  The function has a single input parameter, N, which is the number of doors in the game.

Use the `game` function to estimate the probability that both strategies result in a goat. Let **N=4**.

```{r, eval = FALSE}
require(magrittr)
require(dplyr)

game <- function(N){
  if(N<3) stop("Must have at least 3 doors")
  prize <- sample(c(rep("goat",N-1),"car"), N)
  guess <- sample(1:N,1)
  game <- data.frame(door = 1:N, prize = prize, stringsAsFactors = FALSE) %>% 
    mutate(first_guess = case_when(
      door == guess ~ 1
      , TRUE ~ 0
    )) %>% 
    mutate(potential_reveal = case_when(
        first_guess == 1 ~ 0
      , prize == "car" ~ 0
      , TRUE ~ 1
    )) %>% 
    mutate(reveal = 1*(rank(potential_reveal, ties.method = "random") == 3)) %>% 
    mutate(potential_switch = case_when(
      first_guess == 1 ~ 0
      , reveal == 1 ~ 0
      , TRUE ~ 1
    )) %>% 
    mutate(switch = 1*(rank(potential_switch, ties.method = "random") == 3))
  c(game$prize[game$first_guess == 1], game$prize[game$switch == 1])
}
```

```{r}
vec <- matrix(rep(NA,10000),ncol = 2,byrow = TRUE)
for (i in seq(1:5000)){
  vec[i,] <- game(4)
}
vec <- ifelse(vec=='goat',1,0)

vec_1 <- matrix(rep(NA,10000),ncol = 2,byrow = TRUE)
for (i in seq(5000)){
  if (vec[i,1] == 1 & vec[i,2] == 1){
    vec_1[i,1] <- vec[i,1]
    vec_1[i,2] <- vec[i,2]
  }
}
qwe <- na.omit(vec_1)
length(qwe[,1])/5000


```
So, the probability that both strategies result in a goat when N = 4 is 0.3734

**B**. Continuing from part **C**, what is the probability that at least one of the strategies results in winning a car?

```{r}
# a <- (1/4*1/3)/((1/4*1/3)+(1/4*0)+(1/4*1)+(1/4*1/2))
# b <- (1/4*1/2)/((1/4*1/3)+(1/4*0)+(1/4*1)+(1/4*1/2))
# a+b
vec <- as.data.frame(vec)
vec_1 <- as.data.frame(matrix(rep(NA,10000),ncol = 2,byrow = TRUE))

for (i in seq(5000)){
  if (vec$V1[i] == 0 | vec$V2[i] == 0){
    vec_1$V1[i]<- vec$V1[i]
    vec_1$V2[i] <- vec$V2[i]
  }
}
vec_1<-na.omit(vec_1)
length(vec_1$V1)/5000
```

The probability that at least one of the strategies results in winning a car is 0.6266

**A**. Communicate the precision of your simulated probability in part **B** by calculating a **99\%** confidence interval.

```{r}
length(vec_1$V1)
prop.test(3133,5000,conf.level = 0.99)
```

So, the 99% CI is (0.6087219,0.6441406)

# 2. Probability

Consider a test for a rare genetic condition.  Let T+ denote a test result that indicates the condition is present, while T- denotes absence.  Let D+ and D- denote the true status of the disease.

**C**.  Fill-in the probability table using the following information:

+ P(T+|D+) = .85,  and 
+ P(T-|D-) = .95,  and 
+ P(D+) = 0.001

|        | D+     | D-     |       |
|:---:   |:---:   |:---:   |:---:  |
| T+     | 0.00085| 0.04995| 0.0508|
| T-     | 0.00015|0.94905 |0.9492 |
|        | 0.001  | 0.999  | 1     |

```{r}
0.001*0.85
0.001 - (0.001*0.85)
1-0.001
0.95*0.999
0.999- 0.94905

0.00085 + 0.04995
0.00015 + 0.94905
 0.9492 + 0.0508
```

**B**. Calculate the **negative** predictive value of the test, P(D-|T-).

```{r}
0.94905 / 0.9492
```
 the **negative** predictive value of the test, P(D-|T-) is 0.999842

**A** Create a plot that shows how the **positive** predictive value as a function of the prevalence of disease, P(D+).

```{r, eval = FALSE}
prevalence <- seq(0.001, 0.1, length = 50)
p_plus_d_plus <- NA
for (i in seq(1:50)){
  p_plus_d_plus[i]<- (prevalence[i] * 0.85) + 0.04995
}
plot(prevalence, p_plus_d_plus, xlab = "Prevalence", ylab = "PPV")
```

# 3. Discrete Distributions

Suppose the yearly hospital charges (in thousands of dollars) for a randomly selected Vanderbilt student is a mixture distribution.

For 50% of students, the hospital charges will be $0.  For the remaining 50% of students, the hospital charges are a random variable described by a gamma distribution with shape = 2 and scale = 2.  (Again, in thousands of dollars.)   

```{r}
hospital_charges <- function(N){
  group <- rbinom(N, 1, 0.5)
  charges <- 0*group + rgamma(N, shape = 2, scale = 2)*(1-group)
  charges
}
```

**C**.  What is the 90th percentile for yearly hospital charges for a randomly selected Vanderbilt student?

```{r}
quantile(hospital_charges(1000000),0.90)
```


**B**.  Consider the **class** average yearly hospital charge for the students in a class of size 30.  Plot the density function or a simulated histogram of the class average yearly hospital charge.

```{r}
btw <- NA
for (i in seq(1:50000)){
  a <- hospital_charges(30)
  btw[i]<- mean(a)
}
hist(btw,freq=F)
hist(btw)
```


**A**.  What is the probability that a randomly selected class of size 30 students will have less than 10 students with zero yearly hospital charges?

```{r}
mean(btw)
pbinom(0,9,0.02)
```
the probability that a randomly selected class of size 30 students will have less than 10 students with zero yearly hospital charges is 0.833


# 4. Continuous Distributions

**C.** Suppose diastolic blood pressure (DBP) follows a normal distribution with mean 80 mmHg and SD 15 mmHg. What is the probability that a randomly sampled person’s DBP lies between 70 and 104 mmHg?

```{r}
pnorm(104,80,15) - pnorm(70,80,15)
```


**B.** Suppose a human femur was discovered that is 37 cm long.  Also suppose that using the NHANES data, researchers believe the distribution of femor bones, by sex, are distributed as follows:

+ Female adult femor $\sim N(36, 3.3)$
+ Male adult femor $\sim N(40, 3.4)$



Under the assumption that male and females are equally likely, what is the probability that the discovered femor was from a male?

```{r}
dnorm(x=37,mean=40,sd=3.4)/((dnorm(x=37,mean=36,sd=3.3)+dnorm(x=37,mean=40,sd=3.4)))
```


**A.**  Continuing part **B**, generate a plot of P(femor from male | femor length = x).  Let femor length range from 25 to 50.

```{r}
femor_length <- 25:50
prob_male <- NA
for (i in seq_along(femor_length)){
  prob_male[i]<-dnorm(x=(i+24),mean=40,sd=3.4)/((dnorm(x=(i+24),mean=36,sd=3.3)+dnorm(x=(i+24),mean=40,sd=3.4)))
}
plot.new()
plot.window(xlim = c(25,50), ylim = c(0,1))
lines(femor_length, prob_male)
axis(1)
axis(2)
box()
title(xlab = "Femor Length", ylab = "P( Male | femor length)")
```

# 5. Expectation and Variance

Let us revisit the yearly hospital charges distribution from a previous section.

>**Recall:** The yearly hospital charges (in thousands of dollars) for a randomly selected Vanderbilt student is a mixture distribution. For 50% of students, the hospital charges will be $0.  For the remaining 50% of students, the hospital charges are a random variable described by a gamma distribution with shape = 2 and scale = 2.  (Again, in thousands of dollars.)   

```{r}
hospital_charges <- function(N){
  group <- rbinom(N, 1, 0.5)
  charges <- 0*group + rgamma(N, shape = 2, scale = 2)*(1-group)
  charges
}
```

**C.** What is E[yearly hospital charges]?
```{r}
btw <- NA
for (i in seq(1:50000)){
  a <- hospital_charges(30)
  btw[i]<- mean(a)
}
mean(btw)
```

**B.** Suppose Vanderbilt implements a cap of \$10,000 on yearly student hospital charages.  What is the mean yearly hospital charge under the new policy?

```{r}
btw <- NA
for (i in seq(1:50000)){
  a <- hospital_charges(30)
  btw[i]<- mean(a[a<=10])
}
mean(btw)

```


**A.** What is the variance of yearly hospital charge under the new policy?
```{r}
var(btw)
# Variance decreased under new policy.
```


# 6. Transformations & Sampling Distributions

**C.** Consider the log normal distribution.  If X is a log normal random variable, then log(X) is a normal random variable.  One way to create pseudo-random draws from the log normal distribution is to generate draws from a normal distribution and then to transform the draws by expononentiating.  The parameters of the log normal distribution are the parameters of the underlying normal distribution, $\mu$ and $\sigma$ (or $\sigma^2$).  

Log normal data are prevalent is biological systems and econometrics.

Suppose a blood chemistry measure has a log normal distribution with $\mu$ = 0 and $\sigma$ = 0. Generate an histogram or density curve for the sampling distribution of the median when the sample size is 101.

```{r}
meds <- NA
for (i in seq(1:5000)){
  meds[i] <- rlnorm(101,meanlog = 0,sdlog = 1) %>% median
}
hist(meds)
```


**B.** Below is the CDF function for the kth order statistic when the underlying distribution is log normal with $\mu$ = 0 and $\sigma$ = 1.  Create a plot of the ECDF of the simulated sampling distribution generated in **C** and overlay the CDF using the function below.

```{r}
Fk <- function(x,k,n){
  pbinom(k-1, n, plnorm(x), lower.tail = FALSE)
}
plot(ecdf(meds),col='red')
curve(Fk(x,51,101),add = T,col='blue')
```

**A.** Of the 25th, 50th, and 75th quantiles of the distribution from **B**, which will have the tighest 95% CI?  (Show the sampling distribution of each.)

```{r}
quantile(meds,c(0.25,0.50,0.75))
curve(Fk(x,25,101),add = T,col='blue')
curve(Fk(x,50,101),add = T,col='red')
curve(Fk(x,75,101),add = T,col='black')
```

25th quantile will have the tightest distribution. <br>

# 7. Estimation of CDF and PDF from data

The following code will load the NHANES data and select the first 500 rows.

```{r}
Hmisc::getHdata(nhgh)
d1 <- nhgh[1:500,]
```

**C.** Estimate the distribution of standing height for adult (age > 18) males using the MLE method with a normal distribution.  Create a plot of the estimated density function.

```{r}
d1_male_height <- d1 %>%
  select(sex,age,ht)%>%
  filter(sex == 'male')%>%
  filter(age>18)

dat.mean<- mean(d1_male_height$ht)
dat.sd <- sqrt(((length(d1_male_height$ht)- 1)/length(d1_male_height$ht))*var(d1_male_height$ht))
data.len<-length(d1_male_height$ht)
hist(rnorm(data.len,dat.mean,dat.sd),freq = F)
curve(dnorm(x,dat.mean,dat.sd),add=T,from=0,to=200)

plot(ecdf(rnorm(data.len,dat.mean,dat.sd)))
```


**B.** Estimate the distribution of BMI for adult (age > 18) females using using the method of moment method with the gamma distribution. Create a plot of the estimated density function.

```{r}
d1_female_bmi <- d1 %>% 
  select(sex,age,bmi)%>%
  filter(sex == "female")%>%
  filter(age>18)
# Params
xbar <- mean(d1_female_bmi$bmi)
s2 <- var(d1_female_bmi$bmi)
shape <- (xbar^2) / s2
scale <- s2/xbar

#Method of Moments with gamma dist
Fbmi <- function(x){
  pgamma(x,shape = shape,scale = scale)
}             
  
# Cumalitive Density function 
plot(ecdf(d1_female_bmi$bmi))
curve(Fbmi(x),add = T, col='red')

# Prob Density 
fbmi <- function(x){
  dgamma(x,shape = shape,scale = scale)
}   
hist(d1_female_bmi$bmi,freq = F)
curve(fbmi(x),add=T,col='red')

```


**A.** Estimate the distribution of creatinine (SCr) for adults (age > 18) using the kernel density method with a gaussian kernel.  Create a plot of the estimated density function.

```{r}
d1_scr <- d1%>%
  select(sex,age,SCr)%>%
  filter(age>18)%>%
  filter(!is.na(SCr))

hist(d1_scr$SCr,freq = F)
lines(density(d1_scr$SCr,
        adjust= 1.5,kernel = 'gaussian'))

ecdfstar <- function(t, data, smooth){
  outer(t, data, function(a,b){ pnorm(a, b, smooth)}) %>% rowMeans
}
plot(ecdf(d1_scr$SCr), main = "")
curve(ecdfstar(x,d1_scr$SCr, smooth = 0.1), add = TRUE, lwd = 3, col = "blue")




sample.dist <- NA
    ecdfstar <- function(t, data, smooth=smoo){
    outer(t, data, function(a,b){ pnorm(a, b, smooth)}) %>% rowMeans
    }
    
    tbl <- data.frame(
    x = seq(min(d1_scr$SCr)-sd(d1_scr$SCr),max(d1_scr$SCr) + sd(d1_scr$SCr),by = 0.01)
)
    
tbl$p <- ecdfstar(tbl$x, d1_scr$SCr, smooth=1)
tbl <- tbl[!duplicated(tbl$p),]
qkde <- function(ps, tbl){
  rows <- cut(ps, tbl$p, labels = FALSE)
  tbl[rows, "x"]
}
ps<-runif(100)
Y<-qkde(ps,tbl)
hist(Y,freq=F)
lines(density(d1_scr$SCr,
        adjust= 1.5,kernel = 'gaussian'))

```


# 8. Sample from an estimated distribution

The following code will load the low birth weight data from the MASS package.  The description of the variables in the dataset can be found in the birthwt documentation with the command `?MASS::birthwt`.

```{r}
bwt <- MASS::birthwt
```

**C.** Generate a 95% confidence interval for the mean birthweight of infants whose mothers **did** smoke during pregnancy using the bootstrap.

```{r}
bwt_smoke<- bwt %>%
  select(smoke,bwt)%>%
  filter(smoke==1)

sample_means <- bwt_smoke %>% 
  rep_sample_n(size = length(bwt_smoke$bwt),reps = 5000,replace = T)%>%
  summarise(x_bar = mean(bwt))
quantile(sample_means$x_bar,c(0.025,0.975))
```


**B.** Generate a 95% confidence interval for the mean birthweight of infants whose mothers **did** smoke during pregnancy using the Central Limit Theorem shortcut.

```{r}
clt_mean <- bwt_smoke$bwt%>%mean
clt_sd <- bwt_smoke$bwt%>%sd
lower <- clt_mean - (1.96*clt_sd/sqrt(length(bwt_smoke$bwt)))
lower
upper<-  clt_mean + (1.96*clt_sd/sqrt(length(bwt_smoke$bwt)))
upper
t.test(bwt_smoke$bwt)
```


**A.** Let $\mu_s$ be the mean birthweight of infants whose mothers smoked during pregnancy.  Let $\mu_{ns}$ be the mean for the non-smoking group.  Use simulation to calculate the 95% confidence interval for $\mu_s/\mu_{ns}$.

```{r}
bwt_no_smoke <- bwt%>%
  select(smoke,bwt)%>%
  filter(smoke==0)
out<-NA
for(i in seq(1:100) ){
  mu_s <- rnorm(100,bwt_smoke$bwt%>%mean,bwt_smoke$bwt%>%sd)%>% mean
  mu_ns <- rnorm(100,bwt_no_smoke$bwt%>% mean,bwt_no_smoke$bwt%>%sd) %>% mean
  out[i] <- mu_s/mu_ns
}

quantile(out,c(0.025,0.975))
```


# 9.  Inference

**C.** Suppose two studies were performed looking at the risk of mild complication after hernia repair using open and laparoscopic surgical approaches.  The study results are below.  Using the data from each study individually, perform the hypothesis test that the risk of complication between open and laparoscopic repairs are the same under the usual point null. What is the p-value from each study?  What do you conclude from each study?


| Study 1 | Comp | No comp |
|:---|:---|:---|
| Open | 30 | 70 |
| Lap  | 35 | 65 |

| Study 2 | Comp | No comp |
|:---|:---|:---|
| Open | 600 |     1400 |
| Lap  | 619 |     1381 |


H0 = proportion of open repairs that experience outcome
H0: P0-P1 = 0
H0: P0/P1 = 1
H0: P0/(1-P0) / P1/(1-P1) = 1
H1= proportion of lap experience outcome

```{r}
prop.test(c(30,35),c(100,100))
prop.test(c(600,619),c(2000,2000))
```
 P-value of first study is : 0.5459
 P-value of second study is : 0.5364
 As we can see that P value in both studies is greater than significance point (0.05), so we can say that we accept the null hypothesis in both studies and can say that there is conclusive similarity. But we must also check the threshold point to conclude oue result.

**B.** Suppose that prior to the studies, the researchers established an equivalence threshold of 6 percentage points.  Using the confidence intervals, which studies (if any) showed a conclusive similarity between surgical approaches for the complication rate.  Explain why.

<li>The second study has a conclusive similarity.</li>
<li>Null region = -6 , 6 (equivalence threshold of 6 percentage points)</li>
<li>The ci for the second study is (-3.8%,2.0 %) which is inside of the null region for conclusive similarities, while the first study ci lies inside the null region but not entirely as it is between (-18.9%,8.9%) .</li><br>


**A.** If the data from the studies were combined, what is the smallest equivalence threshold that would identify a conclusive similarity between the surgical approaches?
```{r}
prop.test(c(630,654),c(2100,2100))
```

<li>As CI is (-4%, 1.69%), so it will have equivalence threshold of 4 percentage points ie (-4%,4%) </li><br>

# 10.  Joint Distributions

**C.** Fill in the blank.  The sample correlation is a measure of linear association.<br>

**B.** Explain why predictions from a conditional distribution generally have smaller prediction error than predictions from the marginal distribution.

If two variable are correlated, then conditional distribution is tighter and with less variance than marginal, so we can predict it with less error.
But if two variable are not correlated, then conditional distribution has no advantage than marginal in prediction.<br>

**A.** Use the CLT shortcut to calculate the 95% confidence interval for the correlation of arm circumferance and arm length using the NHANES dataset.  Is the sample correlation a reasonable measure of association for this data?

```{r}
Hmisc::getHdata(nhgh)
cor.test(nhgh$armc,nhgh$arml)
```
As we can see that we get correlation of 0.499 which is generally not considered as good correlation. Also we can see that size of data is 6795 observations which is kind of enough. So, we can say that the sample correlation is not a reasonable measure of association for this data as correlation is not high and we have high number of samples.




